\chapter*{Acknowledgement}

The success and final outcome of the present work required a lot of guidance
and assistance from many people. Without them it would not have been possible
for me to deliver this work and remember the exciting time I had in the past
weeks.

At most I want to thank my parents for their support, that I have been
privileged to benefit from all the years. Though my life involved many
unexpected turnarounds, I never felt left alone and could rely on your
diverse experience whenever the situation required it.

Further I want to thank Prof. Immanuel Bloch and Dr. Monika Aidelsburger for
creating and sustaining a fertile scientific environment. It has been a
refreshing novelty for me to have the opportunity to have inspiring
discussions at almost any time and to seamlessly implement new ideas without
being limited by the present material.

Equally I want to express my gratitude to all personal involved in the
formation process of my thesis. Especially I want to name Christian Schweizer
with whom it had been a great pleasure to work with. Further I want to thank
Till Klostermann for instantly sharing his deep insights in many problems I
have encountered and his help with the optical setup. Hendrik von Raven I want
to thank for his overall guidance with regard to the electronic equipment.
In addition I owe special thanks to Michael Schreiber who helped me understand
the theoretical foundations of atoms in optical lattices, Moritz Berngruber
as a great office colleague and for his help with arbitray physics questions,
David Wei for his hospitality and productive discussions about the
acousto-optic deflectors, and of course biggest thanks to Bodo Hecker for his
advice as electrical engineer and distinct perspective on undergone challenges.
