\chapter{Theory of Optical Lattices}

In this chapter we will recapitulate the theoretical foundations of optical
lattices used in ultracold quantum experiments, thereof we can derive
theoretical requirements for the planned potential generation. Most notably
we are interested in the domain of the sweep duration and resolution that
our setup has to operate in.

\section{General Case}

\subsection{Periodic Potentials}

The Schrödinger equation of a particle state $\ket{\Psi}$ subject to a time
independent one-dimensional potential reads
\begin{align}
  E\ket{\Psi}
  =\hat{H}\ket{\Psi}
  =\left(\frac{\hat{p}^2}{2m}+\hat{V}\right)\ket{\Psi}
  \label{eq:time_independent_schroedinger}.
\end{align}
For a periodic potential $V(x+a)=V(x)$ the Bloch-Theorem \cite{Ashcroft1976}
states that the wavefunction of a particle in a periodic potential can be
written as
\begin{equation}
  \braket{\hat{x}}{\Psi}
  =\Psi(x)
  =e^{iqx}\psi(x)
  \label{eq:bloch_theorem}
\end{equation}
wherein $-\frac{\pi}{a}<q<+\frac{\pi}{a}$ is the quasimomentum, a quantum
number characteristic of the translational symmetry of the periodic potential
\cite[p. 42]{Lewenstein2012} with $a$ being the lattice constant.

The momentum operator is a differential operator in position space
\begin{equation}
  \hat{p}
  =\expval{\hat{p}}{\hat{x}}
  =-i\hbar\dv{x}
  \label{eq:momentum_operator_in_position_space},
\end{equation}
thus we obtain $\hat{p}e^{iqx}=\hbar qe^{iqx}$ and
$\hat{p}^2\ket{\Psi}=e^{iqx}\left(\hat{p}+\hbar q\right)^2\ket{\psi}$ such
that we can rewrite \cref{eq:time_independent_schroedinger} for
the periodic case
\begin{equation}
  E\ket{\psi}
  =\left(\frac{(\hat{p}+\hbar q)^2}{2m}+\hat{V}\right)\ket{\psi}
  =:\hat{H}_{periodic}\ket{\psi}
  \label{eq:periodic_schroedinger}.
\end{equation}

Translational symmetry of the potential enforces discrete particle momenta
$p_n=pn=nk\hbar$ such that we can represent our particle state in a
countable momentum basis
\begin{equation}
  \ket{\psi}
  =\mathbb{1}\ket{\psi}
  =\left(\sum_n\ket{p_n}\bra{p_n}\right)\ket{\psi}
  =\sum_n\braket{p_n}{\psi}\ket{p_n}
  =\sum_nc_n\ket{p_n}
  \label{eq:countable_momentum_state}.
\end{equation}
Inserting this into the right hand side of \cref{eq:periodic_schroedinger}
yields us
\begin{equation}
  \hat{H}_\text{periodic}\ket{\psi}
  =\sum_nc_n\left(\frac{(\hat{p}+\hbar q)^2}{2m}+\hat{V}\right)\ket{p_n}
  =\sum_nc_n\left(\frac{(pn+\hbar q)^2}{2m}+\hat{V}\right)\ket{p_n}
  \label{eq:periodic_hamiltonian_in_momentum_base}
\end{equation}
where we used the eigenvalue equation $\hat{p}\ket{p_n}=pn\ket{p_n}$ to
satisfy the second equal sign. By applying an orthogonal $\bra{p_m}$ to the
left hand side of \cref{eq:periodic_hamiltonian_in_momentum_base} we find
the matrix elements of the periodic Hamiltonian
\begin{equation}
  \hat{H}_{\text{periodic},mn}
  =\mel**{p_m}{\hat{H}_\text{periodic}}{p_n}
  =(m+q/k)^2\frac{\hbar^2k^2}{2m}+\mel**{p_m}{\hat{V}}{p_n}
  \label{eq:periodic_hamiltonian_elements}
\end{equation}
where we used $p=\hbar k$ in the last step.

Usually the potential will be known in position space. Through use of the
completeness relation
\begin{equation}
  \mathbb{1}
  =\int\dd{x}\op{x}{x}
  \label{eq:continous_completeness_relation}
\end{equation}
we are able to transform the potential in momentum space
\begin{equation}
  V(p)
  =\mel**{p_m}{\hat{V}}{p_n}
  =\int\int\dd{x}\dd{y}\braket{p_m}{x}\mel**{x}{\hat{V}}{y}\braket{y}{p_n}
  =\int\frac{\dd{x}}{\sqrt{2\pi\hbar}}e^{ikx(m-n)}V(x)
  \label{eq:potential_fourier_transform}.
\end{equation}
after the identification of $\braket{p}{x}=e^{ipx/\hbar}/\sqrt{2\pi\hbar}$.

\subsection{Harmonic Potentials}

In the former section we discussed the quantum mechancis of a generic
periodic potential. We will now consider the special case of a harmonic
potential of the shape
\begin{equation}
  V(x)
  =V_0\cos^2(kx)
  =\frac{1}{4}V_0\left(e^{+2ikx}+e^{-2ikx}+2\right)
  \label{eq:harmonic_potential}.
\end{equation}
Potentials of such form can be found for example in the standing wave of a
reflected gaussian beam
\begin{equation}
  V(r,x)
  =V_0e^{-2r^2/w(x)^2}\cos^2(kx)
  \approx V_0\cos^2(kx)
  \label{eq:optical_potential}
\end{equation}
wherein $r$ denotes the beam radius, $w(x)$ being spot size parameter and
$k=2\pi/\lambda$ with the laser wavelength $\lambda$. The approximation is
valid for $w(x)\gg r$ which holds in most cases according to
\cite[p.127]{Rom2009}.

Transformation of \cref{eq:harmonic_potential} to momentum space
using \cref{eq:potential_fourier_transform} yields
\begin{equation*}
  V(p)
  =
  \frac{1}{4}V_0
  \int\frac{\dd{x}}{\sqrt{2\pi\hbar}}
  e^{+ipx(m-n)/\hbar}\left(e^{+2ikx}+e^{-2ikx}+2\right)
  =
  \frac{1}{4}V_0\left(2\delta_{m,n}+\delta_{m,n+2}+\delta_{m,n-2}\right).
\end{equation*}
The Hamiltonian \cref{eq:periodic_hamiltonian_elements} of a
harmonic potential then reads
\begin{equation}
  \hat{H}_\text{harmonic}
  =\begin{cases}
    (n+q/k)^2\frac{\hbar^2k^2}{2m}+\frac{1}{2}V_0, & \text{if } n=m\\
    \frac{1}{4}V_0, & \text{if } \abs{n-m}=2\\
    0, & \text{otherwise}
  \end{cases}
  \label{eq:harmonic_hamiltonian}.
\end{equation}
Plotting the eigenvalues of \cref{eq:harmonic_hamiltonian} against
$q$ we find the characteristic energy band structure giving us the allowed
energy ranges.

In the case of optical potentials. We would further introduce the recoil
energy $E_r=(\hbar k)^2/(2m)$ which gives us a natural atom independent scale
to work with.

\section{Free-Particle Approximation}

In the free-particle approximation we consider flat potentials or particles
with higher kinetic than potential energy. In this regime energy gaps
vanish and particles can move freely. For the present work this is not
relevant circumstance as particles as local potentials we are interested
to generate onto the existing lattice structure would just move around
particles.

\section{Tight-Binding Approximation}

In the tight-binding approximation the potentials are considered deep or
particles to have lower kinetic than potential energy, such that interaction
between neighbooring particles is limited. In the optical case potentials can
be easily made deep by increasing the intensity. In this regime energy bands
converge to discrete energy levels with inbetween energies being forbidden.

\subsection{Periodic Potentials}

\subsection{Harmonic Potentials}
